### 2019-12-16 Draft_V5.tex
up

\documentclass[12pt,a4paper,hyperref]{article}
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0, 0.0, 0.55}
	\definecolor{ultramarine}{rgb}{0.07, 0.04, 0.56}
\usepackage{amsmath, natbib, latexsym, array, amssymb,longtable,float, graphicx, appendix,lscape,diagbox,textcomp,placeins}
\usepackage[colorlinks,
            linkcolor=ultramarine,
            anchorcolor=green,
            citecolor=darkblue
            ]{hyperref}
\usepackage[flushleft]{threeparttable}
\usepackage[top=2.7cm, left=3cm, right=3cm, bottom=2.7cm]{geometry}
\usepackage{hyperref}
\newtheorem{myDef}{Definition}
\newtheorem{myTheo}{Theorem}
\newtheorem{myProp}{proposition}
\newtheorem{myRem}{Remark}
\newtheorem{myAssu}{Assumption}
\newtheorem{myCor}{Corollary}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
}

\urlstyle{same}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{pgfplotstable}
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 2, % to 2 places
}

\newenvironment{sequation}{\begin{equation}\tiny}{\end{equation}}
\DeclareMathOperator*{\plim}{plim}
\renewcommand{\floatpagefraction}{0.60}
\renewcommand{\appendixpagename}{\Large Appendix}
\setcounter{secnumdepth}{3}
\begin{document}
\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

\center % Center everything on the page

\HRule \\[0.4cm]
{ \huge \bfseries Dynamic Heterogeneous Panels }\\[0.4cm] % Title of your document
\HRule \\[1.5cm]

\vfill % Fill the rest of the page with whitespace

\end{titlepage}

\newpage
\tableofcontents
\newpage
\section{Model, asymptotic property of LS and IV estimator}
\subsection{The models and Assumptions}
\subsubsection{Fixed effects model}
Consider the dynamic heterogeneous panels data model with fixed effects:
\begin{align}
\begin{split}
y_{i,t}&=\alpha_{i}+\phi_{i} y_{i,t-1}+ \boldsymbol{x}^{'}_{i,t}\boldsymbol{\beta}_{i}+u_{i,t}, \,\, for\,\,i=1,\ldots N;\,t=1,\ldots,T\, , \\
&= \boldsymbol{w}^{'}_{i,t}\boldsymbol{\theta}+ \alpha_{i}+u_{i,t}, \label{1}
\end{split}
\end{align}
where $\boldsymbol{x}_{i,t}$ and $\boldsymbol{\beta}_{i}$ are $k \times 1$ vectors and $\boldsymbol{w}_{i,t}=\left(\phi_{i}, \boldsymbol{\beta}^{'}_{i} \right)^{'}$ is a $\left(1+k\right) \times 1$ vector.
Stacking the $T$ observations for each $i$, we have
\begin{align}
\boldsymbol{y}_{i}=\boldsymbol{W}_{i}\boldsymbol{\theta}_{i}+ \alpha_{i}\boldsymbol{\iota}_{T}+ \boldsymbol{u}_{i}, \label{2}
\end{align}
where $\boldsymbol{y}_{i}=\left(y_{i,1},\ldots, y_{i,T} \right)^{'}$, $\boldsymbol{W}_{i}=\left(\boldsymbol{w}_{i,1},\ldots, \boldsymbol{w}_{i,T} \right)^{'}$, $\iota_{T}=\left(1,\ldots,1 \right)^{'}$ and $\boldsymbol{u}_{i}=\left(u_{i,1}, \ldots, u_{i,T} \right)^{'}$.

Due to the incidental parameters problem arise, we use forward filter to the model by \citet{Moon:2000}, \citet{Hayakawa:2009} and \citet{Hayakawa:2019}. We define the $\left(T-1 \right) \times 1$ forward demeaning matrix as

\begin{align}
\boldsymbol{F}=diag(c_{1}, c_{2}, \ldots c_{T-1})
\begin{bmatrix}
1 & \frac{-1}{T-1} & \cdots & \cdots & \frac{-1}{T-1}\\
\vdots & 1 & \frac{-1}{T-2} & \cdots & \frac{-1}{T-2}\\
\vdots & \vdots & \vdots & \ddots & \vdots \\
0 & 0 & 0 & \cdots & -1 
\end{bmatrix},
\end{align}    
where $c_{t}=\sqrt{\left( T-t\right)\left( T-t+1\right) }$.
Multiply $\boldsymbol{F}$ to model $(\ref{2})$, the model can expressed as  
\begin{align}
\tilde{\boldsymbol{y}}_{i}= \tilde{\boldsymbol{W}}_{i}\boldsymbol{\theta}+\tilde{\boldsymbol{u}}_{i},
\end{align}
where $\tilde{\boldsymbol{y}}_{i}=\boldsymbol{F}\boldsymbol{y}_{i}= \left(\tilde{\boldsymbol{y}}_{i,1}, \ldots, \tilde{\boldsymbol{y}}_{i,T-1} \right)^{'}$,  $\tilde{\boldsymbol{W}}_{i}=\boldsymbol{F}\boldsymbol{W}_{i}= \left(\tilde{\boldsymbol{w}}_{i,1}, \ldots, \tilde{\boldsymbol{w}}_{i,T-1} \right)^{'}$ and $\tilde{\boldsymbol{u}}_{i}=\boldsymbol{F}\boldsymbol{u}_{i}= \left(\tilde{\boldsymbol{u}}_{i,1}, \ldots, \tilde{\boldsymbol{u}}_{i,T-1} \right)^{'}$ with $\tilde{\boldsymbol{y}}_{i,t}=c_{t}\left[y_{i,t}-\left(y_{i,t+1}+\cdots+y_{i,T} \right)/\left(T-t \right) \right]$ for $t=1, \ldots, T-1.$

\subsubsection{Trend model}
Consider the dynamic heterogeneous panels data model with fixed effects and heterogeneous time trends:
\begin{align}
\begin{split}
y_{i,t}&=\alpha_{i}+\eta_{i}t +\phi_{i} y_{i,t-1}+ \boldsymbol{x}^{'}_{i,t}\boldsymbol{\beta}_{i}+u_{i,t}, \,\, for\,\,i=1,\ldots N;\,t=1,\ldots,T\, , \\
&= \boldsymbol{w}^{'}_{i,t}\boldsymbol{\theta}+ \alpha_{i}+\eta_{i}t+u_{i,t}, 
\end{split}
\end{align}
where $\boldsymbol{x}_{i,t}$ and $\boldsymbol{\beta}_{i}$ are $k \times 1$ vectors and $\boldsymbol{w}_{i,t}=\left(\phi_{i}, \boldsymbol{\beta}^{'}_{i} \right)^{'}$ is a $\left(1+k\right) \times 1$ vector.
Stacking the $T$ observations for each $i$, we have
\begin{align}
\boldsymbol{y}_{i}=\boldsymbol{W}_{i}\boldsymbol{\theta}_{i}+ \alpha_{i}\boldsymbol{\iota}_{T}+\eta_{i}\boldsymbol{\tau}_{T} + \boldsymbol{u}_{i}, \label{6}
\end{align}
where $\boldsymbol{\tau}_{T}=\left(1, 2, \ldots, T \right)^{'}$. Again, we define the forward demeaning matrix as

\begin{align}
\boldsymbol{F}^{\tau}=diag(c^{\tau}_{1}, c^{\tau}_{2}, \ldots c^{\tau}_{T-2})
\begin{bmatrix}
1 & \frac{2(-2(T-2))}{(T-1)(T-2)} & \frac{2(-2(T-2)+3)}{(T-1)(T-2)} & \cdots & \frac{2(-2(T-2)+3(T-2))}{(T-1)(T-2)} \\
0 & 1 & \frac{2(-2(T-3))}{(T-2)(T-3)} & \cdots & \frac{2(-2(T-3)+3(T-3))}{(T-3)(T-4)} \\
\vdots &\vdots & \vdots & \ddots  &\vdots \\
0 & \cdots & \cdots & \cdots & \frac{2(-2+3)}{2 \cdot 1}
\end{bmatrix}, \label {7}
\end{align}   
where $c^{\tau}_{t}=\sqrt((T-t)(T-t-1)/(T-t-1)(T-t+2)).$

Multiply $\boldsymbol{F}$ to model $(\ref{6})$, the model can expressed as  
\begin{align}
\tilde{\boldsymbol{y}}^{\tau}_{i}= \tilde{\boldsymbol{W}}^{\tau}_{i}\boldsymbol{\theta}+\tilde{\boldsymbol{u}}^{\tau}_{i},
\end{align}
where $\tilde{\boldsymbol{y}}^{\tau}_{i}=\boldsymbol{F}^{\tau}\boldsymbol{y}_{i}= \left(\tilde{\boldsymbol{y}}^{\tau}_{i,1}, \ldots, \tilde{\boldsymbol{y}}^{\tau}_{i,T-2} \right)^{'}$,  $\tilde{\boldsymbol{W}}^{\tau}_{i}=\boldsymbol{F}^{\tau}\boldsymbol{W}_{i}$ and $\tilde{\boldsymbol{u}}^{\tau}_{i}=\boldsymbol{F}^{\tau}\boldsymbol{u}_{i}.$

\begin{myAssu}
$x_{i,t}$ and $u_{i,t}$ are independently distributed for all $t$ and $s$ .
\end{myAssu}

\subsection{IV estimator}
We use current and lagged values of $\boldsymbol{x}_{i}$ as instruments, as
\begin{align}
\boldsymbol{Z}^{(j)}_{i}=\left( \boldsymbol{x}_{i,\cdot}, \boldsymbol{x}_{i,-1},\ldots, \boldsymbol{x}_{i,-j} \right)^{'},
\end{align}
where $\boldsymbol{Z}^{(j)}_{i}$ is $T \times  (j+1)k$ matrix with $j \geq 1$. \\

Then, we can define IV estimator in fixed effect model of $\boldsymbol{\theta}^{b}_{i}$ as
\begin{align}
\hat{\boldsymbol{\theta}}^{b}_{IV,i}= \left( \tilde{\boldsymbol{W}}^{'}_{i}\boldsymbol{P}^{(j)}_{i} \tilde{\boldsymbol{W}}_{i} \right)^{-1} \tilde{\boldsymbol{W}}^{'}_{i}\boldsymbol{P}^{(j)}_{i} \tilde{\boldsymbol{y}}_{i},
\end{align}
where $\boldsymbol{P}^{(j)}_{i}=\boldsymbol{Z}^{(j)}_{i} \left(\boldsymbol{Z}^{(j)'}_{i}\boldsymbol{Z}^{(j)}_{i} \right)^{-1}\boldsymbol{Z}^{(j)'}_{i}$ and $b=1$ corresponds to the fixed effects model while $b=2$ corresponds to the trend models.

The

\subsection{Asymptotic property of LS estimator}

Based on heterogenous dynamic panel data model $(\ref{1})$, we can obtain fixed effect estimator as
\begin{align}
\hat{\boldsymbol{\theta}}_{LS,i}=
\begin{pmatrix}
\hat{\phi}_{i} \\
\hat{\beta}_{i}
\end{pmatrix}=
\begin{pmatrix}
\frac{\sum^{T}_{t=1}\tilde{y}^{2}_{i,t-1}}{T} & \frac{\sum^{T}_{t=1}\tilde{y}_{i,t-1}\tilde{x}_{i,t}}{T} \\
\frac{\sum^{T}_{t=1}\tilde{x}_{i,t}\tilde{y}_{i,t-1}}{T} &  \frac{\sum^{T}_{t=1}\tilde{x}_{i,t}^{2}}{T}
\end{pmatrix}^{-1}
\begin{pmatrix}
\frac{\sum^{T}_{t=1}\tilde{y}_{i,t-1}\tilde{y}_{i,t}}{T} \\
\frac{\sum^{T}_{t=1}\tilde{x}_{i,t}\tilde{y}_{i,t}}{T}
\end{pmatrix},
\end{align}

where $\tilde{y}_{i,t}=y_{i,t}-\bar{y}_{i}$, $\tilde{y}_{i,t-1}=y_{i,t-1}-\bar{y}_{i,-1}\iota_{T}$ and $\tilde{x}_{i,t}=x_{i,t}-\bar{x}_{i}$ with $\bar{y}_{i}=\frac{1}{T}\sum^{T}_{t=1}y_{i,t}$, $\bar{y}_{i,-1}=\frac{1}{T}\sum^{T}_{t=1}y_{i,t-1}$,  $\bar{x}_{i}=\frac{1}{T}\sum^{T}_{t=1}x_{i,t}$.  Under equation $(\ref{1})$, we have
\begin{align}
\begin{pmatrix}
\hat{\phi}_{i}-\phi_{i} \\
\hat{\beta}_{i}-\beta_{i}
\end{pmatrix}=
\begin{pmatrix}
\frac{\sum^{T}_{t=1}\tilde{y}^{2}_{i,t-1}}{T} & \frac{\sum^{T}_{t=1}\tilde{y}_{i,t-1}\tilde{x}_{i,t}}{T} \\
\frac{\sum^{T}_{t=1}\tilde{x}_{i,t}\tilde{y}_{i,t-1}}{T} &  \frac{\sum^{T}_{t=1}\tilde{x}_{i,t}^{2}}{T}
\end{pmatrix}^{-1}
\begin{pmatrix}
\frac{\sum^{T}_{t=1}\tilde{y}_{i,t-1}\tilde{u}_{i,t}}{T} \\
\frac{\sum^{T}_{t=1}\tilde{x}_{i,t}\tilde{u}_{i,t}}{T}
\end{pmatrix},
\end{align}
where $\tilde{u}_{i,t}$ is $u_{i,t}-\bar{u}_{i}$ with $\bar{u}_{i}=\frac{1}{T}\sum^{T}_{t=1}u_{i,t}.$

Now, we can investigate asymptotic bias by taking the probability limit as
\begin{align}
A^{(1)}_{\phi i}=\plim_{T  \rightarrow \infty}\left( \frac{\sum^{T}_{t=1}\tilde{y}_{i,t-1}\tilde{u}_{i,t}}{T} \right). \label{11}
\end{align}
Then $A^{(1)}_{i}$ can be taken expectations as
\begin{align}
\begin{split}
A^{(1)}_{\phi i}&=E\left(y_{i,t-1}-\bar{y}_{i,-1} \right) \left(u_{i,t}-\bar{u}_{i} \right) \\
&= E\left( y_{i,t-1}u_{i,t}\right) - E\left( y_{i,t-1} \bar{u}_{i}  \right)-E\left(\bar{y}_{i,-1}   u_{i,t}\right)+E\left(\bar{y}_{i,-1}  \bar{u}_{i}\right),
\end{split}
\end{align}
where $E\left( y_{i,t-1}u_{i,t}\right)=0$.

And we assume $y_{i,t}$ has started from a long time period in the past, so we have
\begin{align}
y_{i,t}=\frac{\alpha_{i}}{\left(1-\phi_{i} \right)}+\sum^{\infty}_{s=0}\beta_{i}\phi^{s}_{i}x_{i,t -s}+\sum^{\infty}_{s=0}\phi^{s}_{i}u_{i,t-s},
\end{align}

Then, we have
\begin{align}
\begin{split}
A^{(1)}_{\phi i}&=-E\left( \left(\sum^{\infty}_{s=0}\phi^{s}_{i}u_{i,t-s-1} \right) \left(\frac{1}{T}\sum^{T}_{t=1}u_{i,t} \right)  \right)-E\left(\frac{u_{i,t}}{T}\sum^{T}_{t=1}\sum^{\infty}_{s=0}\phi^{s}_{i}u_{i,t-s-1} \right)+ \\
&\left(\frac{1}{T}\sum^{T}_{t=1}\sum^{\infty}_{s=0}\phi^{s}_{i}u_{i,t-s-1} \right)\left( \frac{1}{T}\sum^{T}_{t=1}u_{i,t} \right). \label{12}
\end{split}
\end{align}
 Hence, from above equation, we have
 \begin{equation}
\begin{split}
A^{(1)}_{\phi i}=&-\dfrac{1}{T} E\left\{  \left( u_{i,t-1}+u_{i,t-2}\phi_{i}^{1}+u_{i,t-3}\phi_{i}^{2}+\ldots\right)\left(u_{i,1}+\ldots+u_{i,t-1}+u_{i,t}+\ldots+u_{i,T}\right) \right\} -\\
 &\dfrac{1}{T} E \left\{  u_{i,t}\sum^{T}_{s=1}\left( u_{i,s-1}\phi_{i}^{0}+u_{i,s-2}\phi_{i}^{1}+\ldots+u_{i,s-t-1}\phi_{i}^{t}+\ldots+u_{i,s-T-1}\phi_{i}^{T}+\ldots \right)  \right\}  + \\
 +& \dfrac{1}{T} E\{ \left( \sum^{T}_{s=1}u_{i,s-1}\phi_{i}^{0}+\sum^{T}_{s=1}u_{i,s-2}\phi_{i}^{1}+\ldots+\sum^{T}_{s=1}u_{i,s-t-1}\phi_{i}^{t}+\ldots
+\sum^{T}_{s=1}u_{is-T-1}\phi_{i}^{T}+\ldots \right)\\
 &\left(\dfrac{1}{T}\sum^{T}_{s=1}u_{i,s}\right) \} \\
 =&-\dfrac{\sigma_{u}^{2}}{T}\dfrac{(1-\phi_{i}^{t-1})}{1-\phi_{i}}-\dfrac{\sigma^{2}_{u}}{T}\dfrac{(1-\phi_{i}^{T-t})}{(1-\phi_{i})}+\dfrac{\sigma^{2}_{u}}{T} \left( \dfrac{1}{1-\phi_{i}}-\dfrac{1}{T}\dfrac{(1-\phi_{i}^{T})}{(1-\phi_{i})^{2}}\right) \\
 =&-\dfrac{\sigma_{u}^{2}}{T(1-\phi_{i})}\left( 1-\phi_{i}^{t-1}-\phi_{i}^{T-t}+\dfrac{1}{T}\dfrac{(1-\phi_{i}^{T})}{(1-\phi_{i})}  \right).
 \end{split}
\end{equation}

Therefore, we can see the bias of $\hat{\phi}_{i}$ is $O\left(T^{-1}\right)$.

To be more compact ,we can rewrite the model as,
\begin{align}
\tilde{\boldsymbol{y}}_{i}=\tilde{\boldsymbol{W}}_{i}\boldsymbol{\theta}_{i}+\tilde{\boldsymbol{u}}_{i},
\end{align}
where $\tilde{\boldsymbol{y}}_{i}=\left(\tilde{\boldsymbol{y}}_{i,1},\ldots, \tilde{\boldsymbol{y}}_{i,T}  \right)^{'}$ is $T \times 1$ vector,  $\tilde{\boldsymbol{W}}_{i}=\left( \tilde{\boldsymbol{w}}_{i,1},\ldots, \tilde{\boldsymbol{w}}_{i,T} \right)^{'}$ is $T \times 2$ matrix and $\tilde{\boldsymbol{u}}_{i}=\left(\tilde{\boldsymbol{u}}_{i,1}, \ldots, \tilde{\boldsymbol{u}}_{i,T} \right)$ is $T \times 1$ vector with $\tilde{\boldsymbol{w}}_{i,t}=\left(y_{i,t-1}-\bar{y}_{i,-1}, x_{i,t}-\bar{x}_{i}  \right)^{'}$, for $t=1,\ldots, T$ and $i=1,\ldots, N.$

Also, we define our interested parameter as
\begin{align}
\left(\phi_{i},\beta_{i}  \right)^{'}=\boldsymbol{\theta}_{i}=\boldsymbol{\theta}+\boldsymbol{\lambda}_{i},
\end{align}
where $\boldsymbol{\lambda}_{i}\overset{i.i.d.}{\sim} \left(\boldsymbol{0}, \boldsymbol{\Sigma}_{\lambda} \right).$
The lest square estimator, $\hat{\boldsymbol{\theta}}_{LS,i}$, can be expressed as
\begin{align}
\hat{\boldsymbol{\theta}}_{LS,i}=\left(\frac{\tilde{\boldsymbol{W}}^{'}_{i} \tilde{\boldsymbol{W}}_{i}}{T}  \right)^{-1} \frac{ \tilde{\boldsymbol{W}}^{'}_{i}\tilde{\boldsymbol{y}}_{i}}{T}. \label{13}
\end{align}

From above discussion and assumptions, we have following theorem \\
\begin{myTheo}
\begin{align}
\sqrt{T}\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i}\right) \overset{d}{\to} N \left(\boldsymbol{0}, \boldsymbol{Q}_{i}^{-1}\boldsymbol{\Sigma}_{LS,i}\boldsymbol{Q}_{i}^{-1}  \right),
\end{align}
where $\boldsymbol{\Sigma}_{LS,i}=\plim_{T \to \infty} T^{-1}\tilde{\boldsymbol{W}}^{'}_{i} \tilde{\boldsymbol{u}}_{i} \tilde{\boldsymbol{u}}^{'}_{i}  \tilde{\boldsymbol{W}}_{i} $ and $\boldsymbol{Q}_{i}=\plim_{T \to \infty} T^{-1}\tilde{\boldsymbol{W}}^{'}_{i}\tilde{\boldsymbol{W}}_{i}$
\end{myTheo}

\subsubsection{Mean group LS estimator }
Now, we define the mean group estimator of $\boldsymbol{\theta}$:
\begin{align}
\hat{\boldsymbol{\theta}}_{LSMG}=\frac{1}{N}\sum^{N}_{i=1}\hat{\boldsymbol{\theta}}_{LSi}.
\end{align}
And we can show that the asymptotic property of $\hat{\boldsymbol{\theta}}_{LSMG}$, as
\begin{align}
\begin{split}
\hat{\boldsymbol{\theta}}_{LSMG}&=N^{-1}\sum^{N}_{i=1}\left(\frac{\tilde{\boldsymbol{W}}^{'}_{i} \tilde{\boldsymbol{W}}_{i}}{T}  \right)^{-1} \frac{ \tilde{\boldsymbol{W}}^{'}_{i}\tilde{\boldsymbol{y}}_{i}}{T} \\
&=\bar{\boldsymbol{\theta}}+N^{-1}\sum^{N}_{i=1}\left(\frac{\tilde{\boldsymbol{W}}^{'}_{i} \tilde{\boldsymbol{W}}_{i}}{T}  \right)^{-1} \frac{ \tilde{\boldsymbol{W}}^{'}_{i}\tilde{\boldsymbol{u}}_{i}}{T},
\end{split}
\end{align}
where $\bar{\boldsymbol{\theta}}=N^{-1}\sum^{N}_{i=1}\boldsymbol{\theta}_{i}.$
For fixed $N$ and large $T$, we have
\begin{align}
\plim_{T \to \infty}\hat{\boldsymbol{\theta}}_{LSMG} =  \bar{\boldsymbol{\theta}}+ N^{-1}\sum^{N}_{i=1}\plim_{T \to \infty}\left(\frac{\tilde{\boldsymbol{W}}^{'}_{i} \tilde{\boldsymbol{W}}_{i}}{T}  \right)^{-1}\plim_{T \to \infty}\left( \frac{ \tilde{\boldsymbol{W}}^{'}_{i}\tilde{\boldsymbol{u}}_{i}}{T}\right)
\end{align}
 Then, from section $1.1$, we know that $\plim_{T \to \infty}\left( \frac{ \tilde{\boldsymbol{W}}^{'}_{i}\tilde{\boldsymbol{u}}_{i}}{T}\right)=O_{p}(1)$. Thus,  we can obtain
\begin{align}
\plim_{T \to \infty}\hat{\boldsymbol{\theta}}_{LSMG} =  \bar{\boldsymbol{\theta}}.
\end{align}
When $N \to \infty$ and $T \to \infty$ and by the law of large numbers, we can see that
\begin{align}
\plim_{T \to \infty, N \to \infty}\hat{\boldsymbol{\theta}}_{LSMG}=\boldsymbol{\theta}.
\end{align}

And the variance estimator of $\hat{\boldsymbol{\theta}}_{LSMG}$ is given by
\begin{align}
\hat{\boldsymbol{\Sigma}}_{LS,\lambda}=\frac{1}{N-1}\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)^{'}.
\end{align}

We show that $\hat{\boldsymbol{\Sigma}}_{LS,\lambda}$ is consistent when $N \to \infty$ and $T \to \infty.$
\begin{align}
\begin{split}
&\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)^{'}=\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LS,i}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right) \right)\left( \hat{\boldsymbol{\theta}}_{LS,i}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right)\right)^{'}\\
&+\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LSMG}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right)\right) \left( \hat{\boldsymbol{\theta}}_{LSMG}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right)\right)^{'}\\
&-\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LSMG}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right)\right) \left( \hat{\boldsymbol{\theta}}_{LS,i}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right)\right)^{'} \\
&-\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LS,i}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right)\right) \left( \hat{\boldsymbol{\theta}}_{LSMG}- E\left(\hat{\boldsymbol{\theta}}_{LS, i}\right)\right)^{'}. \label{20}
\end{split}
\end{align}

Taking expectation on equation $(\ref{20})$, we have
\begin{align}
\begin{split}
&E\left( \sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)^{'} \right)=
 \sum^{N}_{i=1}Var\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)+
\sum^{N}_{i=1}\hat{\boldsymbol{\theta}}_{LSMG}\hat{\boldsymbol{\theta}}^{'}_{LSMG}+  \\
&\sum^{N}_{i=1}E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right) E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)^{'}-
\sum^{N}_{i=1}\hat{\boldsymbol{\theta}}_{LSMG}\hat{\boldsymbol{\theta}}^{'}_{LS,i}+
\sum^{N}_{i=1}E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)\hat{\boldsymbol{\theta}}^{'}_{LS,i}-
\sum^{N}_{i=1}\hat{\boldsymbol{\theta}}_{LSMG}E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)^{'}- \\
&\sum^{N}_{i=1}\hat{\boldsymbol{\theta}}_{LS,i}\hat{\boldsymbol{\theta}}^{'}_{LSMG}+
\sum^{N}_{i=1}\hat{\boldsymbol{\theta}}_{LS,i}E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)^{'}-
\sum^{N}_{i=1}E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right) E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)^{'} = \\
&\sum^{N}_{i=1}Var\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)-N E \left(\hat{\boldsymbol{\theta}}_{LSMG}\hat{\boldsymbol{\theta}}^{'}_{LSMG} \right)+\sum^{N}_{i=1}E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right) E\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)^{'}
\end{split}
\end{align}
and
\begin{align}
E \left(\hat{\boldsymbol{\theta}}_{LSMG}\hat{\boldsymbol{\theta}}^{'}_{LSMG} \right)=\frac{1}{N^{2}} \sum^{N}_{i=1}\sum^{N}_{j=1}E \left(\hat{\boldsymbol{\theta}}_{LS,i}\hat{\boldsymbol{\theta}}^{'}_{LS,i} \right).
\end{align}

And, we also have
\begin{align}
E \left(\hat{\boldsymbol{\theta}}_{LSMG}\hat{\boldsymbol{\theta}}^{'}_{LSMG} \right)=\frac{1}{N^{2}}\left( \sum^{N}_{i=1}Var\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)+ \sum^{N}_{i=1}\sum^{N}_{j=1}E \left(\hat{\boldsymbol{\theta}}_{LS,i}\hat{\boldsymbol{\theta}}^{'}_{LS,i} \right) \right).
\end{align}
Then,
\begin{align}
\begin{split}
&E\left( \sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)^{'} \right)= \\
&\left(1-\frac{1}{N} \right)\sum^{N}_{i=1}Var\left(\hat{\boldsymbol{\theta}}_{LS,i}\right)+ \sum^{N}_{i=1}E \left(\hat{\boldsymbol{\theta}}_{LS,i}\hat{\boldsymbol{\theta}}^{'}_{LS,i} \right)- \frac{1}{N}\sum^{N}_{i=1}\sum^{N}_{j=1}E \left(\hat{\boldsymbol{\theta}}_{LS,i}\hat{\boldsymbol{\theta}}^{'}_{LS,i} \right). \label{24}
\end{split}
\end{align}
From above equation $(\ref{24})$, we can observe the bias term as,
\begin{align}
\aleph=\sum^{N}_{i=1}E \left(\hat{\boldsymbol{\theta}}_{LS,i}\hat{\boldsymbol{\theta}}^{'}_{LS,i} \right)- \frac{1}{N}\sum^{N}_{i=1}\sum^{N}_{j=1}E \left(\hat{\boldsymbol{\theta}}_{LS,i}\hat{\boldsymbol{\theta}}^{'}_{LS,i} \right). \label{25}
\end{align}

 Taking expectation on equation $(26)$, we have
\begin{align}
E\left( \hat{\boldsymbol{\theta}}_{LS,i} \right)=\boldsymbol{\theta}+E \left(b_{i} \right),
\end{align}
From equation $(\ref{25})$, we know that
 \begin{align}
\aleph=\sum^{N}_{i=1}E \left(b_{i}\right) E\left(b^{'}_{i} \right)- \frac{1}{N}\sum^{N}_{i=1}\sum^{N}_{j=1}E \left(b_{i}\right) E\left(b^{'}_{j} \right).
 \end{align}
When $T \to \infty$, $\aleph=0$.
Therefore, we have following theorem \\

\begin{myTheo}
When $(T,N) \overset{j}{\to} \infty$ such that $N/T \to c$ with $0<c< \infty$,
\begin{enumerate}
\item \begin{align}
\sqrt{N}\left(\hat{\boldsymbol{\theta}}_{LSMG}-\boldsymbol{\theta}  \right)\overset{d}{\to} N\left(\boldsymbol{0},\boldsymbol{\Sigma}_{LS,\lambda} \right).
\end{align}
\item \begin{align}
\hat{\boldsymbol{\Sigma}}_{LS,\lambda} \overset{p}{\to}  \boldsymbol{\Sigma}_{LS,\lambda}
\end{align}
where
\begin{align}
 \hat{\boldsymbol{\Sigma}}_{LS,\lambda}=\frac{1}{N-1}\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)\left( \hat{\boldsymbol{\theta}}_{LS,i}- \hat{\boldsymbol{\theta}}_{LSMG}\right)^{'}.
\end{align}
\end{enumerate}
\end{myTheo}

\subsection{Asymptotic property of IV estimator}
We use current and lagged values of $\boldsymbol{x}_{i}$ as instruments, as
\begin{align}
\tilde{\boldsymbol{Z}}_{i,t}=\left( \tilde{\boldsymbol{x}}_{i}, \tilde{\boldsymbol{x}}_{i,-1} \right)^{'},
\end{align}
where $\tilde{\boldsymbol{Z}}_{i}$ is $T \times  2$ vector. \\

\begin{myAssu}
$\boldsymbol{A}_{i}=\plim_{T \to \infty}\tilde{\boldsymbol{A}}_{i,T}$ has full column rank, $\boldsymbol{B}_{i}=\plim_{T \to \infty}\tilde{\boldsymbol{B}}_{i,T}$ and $\boldsymbol{\Sigma}_{i}=\plim_{T \to \infty} T^{-1}\boldsymbol{Z}^{'}_{i}\boldsymbol{u}_{i} \boldsymbol{u}_{i}^{'}\boldsymbol{Z}$ has positive definite, uniformly, where $\tilde{\boldsymbol{A}}_{i,T}= \frac{1}{T}\tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{W}}_{i}$ and $\tilde{\boldsymbol{B}}_{i,T}= \frac{1}{T}\tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{Z}}_{i}$ \\
\end{myAssu}

Then, the IV estimator can be expressed as
\begin{align}
\boldsymbol{\hat{\theta}}_{IV,i}=\left(\tilde{\boldsymbol{A}}_{i,T}^{'}\tilde{\boldsymbol{B}}_{i,T}^{-1}\tilde{\boldsymbol{A}}_{i,T} \right)\tilde{\boldsymbol{A}}_{i,T}^{'}\tilde{\boldsymbol{B}}_{i,T}^{-1}\tilde{\boldsymbol{g}}_{i,T},
\end{align}
where
\begin{align}
\tilde{\boldsymbol{g}}_{i,T}=\frac{1}{T}\tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{y}}_{i},
\end{align}
 and $\tilde{\boldsymbol{W}}_{i}=\left( \tilde{w}^{'}_{i,1},\ldots, \tilde{w}^{'}_{i,T} \right)^{'}$ is $T \times 2$ matrix

From above equation, we have
\begin{align}
\sqrt{T}\left(\boldsymbol{\hat{\theta}}_{IV,i}-\boldsymbol{\theta}_{i} \right)=\left(\tilde{\boldsymbol{A}}_{i,T}^{'}\tilde{\boldsymbol{B}}_{i,T}^{-1}\tilde{\boldsymbol{A}}_{i,T} \right)\tilde{\boldsymbol{A}}_{i,T}^{'}\tilde{\boldsymbol{B}}_{i,T}^{-1}\left(T^{-1/2}  \tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{u}}_{i} \right)
\end{align}
Then, the property of $T^{-1/2}  \tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{u}}_{i} $ is given by following proposition.\\
\begin{myProp}
Under above assumptions, as $\left(N, T \right)\overset{j}{\to} \infty$ such that $N/T \to c$ with $0 < c<\infty $, for each $i$, we have
\begin{align}
\begin{split}
&N^{-1} \tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{u}}_{i}\overset{p}{\to} \boldsymbol{0},  \\
&\,\,\,\,and\,\,\,\,\, \\
&T^{-1/2}  \tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{u}}_{i} \overset{d}{\to} N\left(\boldsymbol{0},\boldsymbol{\Sigma}_{i} \right).
\end{split}
\end{align}
\end{myProp}

Thus, IV  estimator, $\boldsymbol{\theta}_{IV, i}$ is $\sqrt{T}$ consistent to $\boldsymbol{\theta}_{i}$ and this estimator does not have Nickell's bias. Then, we have following theorem\\
$\mathbf{Theorem \,\,1}$ \\
As $\left(N,T \right)\to \infty $ such that $N/T \to c$ with $0< c< \infty$. for each $i$,
\begin{align}
\sqrt{T}\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i}  \right)\overset{d}{\to} N \left(\boldsymbol{0},\,\left(\boldsymbol{A}_{i}^{'}\boldsymbol{B}_{i}^{-1}\boldsymbol{A}_{i} \right)^{-1}\boldsymbol{A}_{i}^{'}\boldsymbol{B}_{i}^{-1}   \boldsymbol{\Sigma}_{i}\boldsymbol{B}_{i}^{-1} \boldsymbol{A}_{i}\left(\boldsymbol{A}_{i}^{'}\boldsymbol{B}_{i}^{-1}\boldsymbol{A}_{i} \right)\right).
\end{align}

\subsubsection{Mean group IV estimator }
Now, we define the mean group estimator of $\boldsymbol{\theta}$:
\begin{align}
\hat{\boldsymbol{\theta}}_{IVMG}=\frac{1}{N}\sum^{N}_{i=1}\hat{\boldsymbol{\theta}}_{IVi}.
\end{align}
And we can show that the asymptotic property of $\hat{\boldsymbol{\theta}}_{IVMG}$, as
\begin{align}
\hat{\boldsymbol{\theta}}_{IVMG}-\boldsymbol{\theta}=\frac{1}{N}\sum^{N}_{i=1}\left(\hat{\boldsymbol{\theta}}_{IV,i}- \boldsymbol{\theta}\right)=\frac{1}{N}\sum^{N}_{i=1}\left(\hat{\boldsymbol{\theta}}_{IV,i}- \boldsymbol{\theta}_{i}\right)+\frac{1}{N}\sum^{N}_{i=1}\boldsymbol{\lambda}_{i},
\end{align}
where
\begin{align}
\begin{split}
\frac{1}{N}\sum^{N}_{i=1}\left(\hat{\boldsymbol{\theta}}_{IV,i}- \boldsymbol{\theta}_{i}\right)&=\frac{1}{N}\sum^{N}_{i=1} \left(\tilde{\boldsymbol{A}}_{i,T}^{'}\tilde{\boldsymbol{B}}_{i,T}^{-1}\tilde{\boldsymbol{A}}_{i,T} \right)^{-1}\tilde{\boldsymbol{A}}_{i,T}^{'}\tilde{\boldsymbol{B}}_{i,T}^{-1}\left(T^{-1}  \tilde{\boldsymbol{Z}}^{'}_{i}\tilde{\boldsymbol{u}}_{i} \right) \\
&= O_{p}\left(\delta^{-2}_{NT} \right),
\end{split}
\end{align}
where $\delta_{NT}=\min \{\sqrt{N},\,\sqrt{T} \}$.
We note that $\frac{1}{\sqrt{N}}\sum^{N}_{i=1}\boldsymbol{\lambda}_{i}=O_{p}\left(N^{-1/2} \right)$, if $\delta_{NT}^{-(2+\varsigma)}\to 0$ for any $\varsigma >0$, we have
\begin{align}
\sqrt{N}\left(\hat{\boldsymbol{\theta}}_{IVMG}-\boldsymbol{\theta}  \right)=\frac{1}{\sqrt{N}}\sum^{N}_{i=1}\boldsymbol{\lambda}_{i}+o_{p} \left(1 \right).
\end{align}

And the variance estimator of $\hat{\boldsymbol{\theta}}_{IVMG}$ is given by
\begin{align}
\hat{\boldsymbol{\Sigma}}_{IV,\lambda}=\frac{1}{N-1}\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{IV,i}- \hat{\boldsymbol{\theta}}_{IVMG}\right)\left( \hat{\boldsymbol{\theta}}_{IV,i}- \hat{\boldsymbol{\theta}}_{IVMG}\right)^{'}. \label{44}
\end{align}
Follow \citet{Norkute:2019}, we can show that $\hat{\boldsymbol{\Sigma}}_{IV,\lambda}$ is consistent and it does not have small $T$ bias.
Firstly, we decompose $(\ref{44})$ as
\begin{align}
\begin{split}
&\sum^{N}_{i=1}\left( \hat{\boldsymbol{\theta}}_{IV,i}- \boldsymbol{\theta}+\boldsymbol{\theta}-    \hat{\boldsymbol{\theta}}_{IVMG}\right)\left( \hat{\boldsymbol{\theta}}_{IV,i}- \boldsymbol{\theta}+\boldsymbol{\theta}-    \hat{\boldsymbol{\theta}}_{IVMG}\right)^{'}= \\
& \sum^{N}_{i=1} \boldsymbol{\lambda}_{i}\boldsymbol{\lambda}^{'}_{i}+\sum^{N}_{i=1}\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i} \right)\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i} \right)^{'}+\sum^{N}_{i=1}\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i}  \right) \boldsymbol{\lambda}_{i}+\sum^{N}_{i=1}\boldsymbol{\lambda}_{i} \left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i}  \right) -\\
&N\left(\boldsymbol{\theta}-\hat{\boldsymbol{\theta}}_{IVMG} \right)^{'}\left(\boldsymbol{\theta}-\hat{\boldsymbol{\theta}}_{IVMG} \right). \label{45}
\end{split}
\end{align}
Then we can show consistent of $\hat{\boldsymbol{\Sigma}}_{IV,\lambda}$ as
\begin{align}
\begin{split}
&\hat{\boldsymbol{\Sigma}}_{IV,\lambda}-\boldsymbol{\Sigma}_{IV,\lambda}=
 \frac{1}{N-1}\sum^{N}_{i=1}\left( \boldsymbol{\lambda}_{i}\boldsymbol{\lambda}^{'}_{i}-\boldsymbol{\Sigma}_{IV,\lambda}\right) +\frac{1}{N-1}\sum^{N}_{i=1}\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i} \right)\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i} \right)^{'} \\
&+\frac{1}{N-1}\sum^{N}_{i=1}\left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i}  \right) \boldsymbol{\lambda}_{i}+\frac{1}{N-1}\sum^{N}_{i=1}\boldsymbol{\lambda}_{i} \left(\hat{\boldsymbol{\theta}}_{IV,i}-\boldsymbol{\theta}_{i}  \right) -\\
&\frac{N}{N-1}\left(\boldsymbol{\theta}-\hat{\boldsymbol{\theta}}_{IVMG} \right)^{'}\left(\boldsymbol{\theta}-\hat{\boldsymbol{\theta}}_{IVMG} \right)  =o_{p}(1).
\end{split}
\end{align}

Then, we can see that the asymptotic property of $\hat{\boldsymbol{\theta}}_{IVMG} $ as,
\begin{align}
\sqrt{N}\left(\hat{\boldsymbol{\theta}}_{IVMG}-\boldsymbol{\theta}  \right)\overset{d}{\to} N\left(\boldsymbol{0},\boldsymbol{\Sigma}_{IV,\lambda} \right).
\end{align}

\section{Choosing the number of instruments}
When $\{ x_{i,t} \}^{T}_{t=1-j}$ for $j \geq 1$ are available $\left( j+1 \right)$ instruments can be used. And we also know that using more instruments, estimator are more efficient but more biased.

\section{Estimation method on dynamic heterogeneous panel data model with  multifactor error structure}
For convenient, we assume the number of regressor is $1$ and we express the model as
\begin{align}
y_{i,t}=\phi_{i} y_{i,t-1}+ \sum^{k}_{\ell=1}\beta_{\ell i}x_{\ell i,t}+u_{i,t}, \,\, for\,\,i=1,\ldots N;\,t=1,\ldots,T, \, \ell=1,\ldots, k. \label{46}
\end{align}
Consider the model $(\ref{45})$, we drawn $x_{\ell i,t}$ as
\begin{align}
x_{\ell i,t}=\boldsymbol{\gamma}^{0'}_{xi}\boldsymbol{f}^{0}_{xt}+\varepsilon_{xi,t}
\end{align}
and the idiosyncratic errors of the process for $y_{i,t}$ as
\begin{align}
u_{i,t}=\boldsymbol{\gamma}^{0'}_{yi}\boldsymbol{f}^{0}_{yt}+\varepsilon_{yi,t},
\end{align}
where $\boldsymbol{\gamma}^{0}_{yi}$ and $\boldsymbol{\gamma}^{0}_{xi}$ are $m_{y}\times 1$ and $m_{x}\times 1$ true
 factor loading respectively, $\boldsymbol{f}^{0}_{yt}$  and  $\boldsymbol{f}^{0}_{xt}$ are  $m_{y}\times 1$ and $m_{x}\times 1$ true vector of unobservable factors respectively.
\subsection{Norkutes' (2019) IVMG estimator}
 We asymptotically eliminate the common factor in $\boldsymbol{x}_{i}$ by projecting matrix, $\boldsymbol{M}_{F^{0}_{x}}$.
 \begin{align}
 \boldsymbol{M}_{F^{0}_{x}}=\boldsymbol{I}_{T}-\boldsymbol{F}^{0}_{x}\left(\boldsymbol{F}^{0'}_{x}\boldsymbol{F}^{0}_{x}  \right)^{-1}\boldsymbol{F}^{0'}_{x} ; \boldsymbol{M}_{F^{0}_{x,-1}}=\boldsymbol{I}_{T}-\boldsymbol{F}^{0}_{x,-1}\left(\boldsymbol{F}^{0'}_{x,-1}\boldsymbol{F}^{0}_{x,-1}  \right)^{-1}\boldsymbol{F}^{0'}_{x,-1}
 \end{align}
And using the defactored covariates as instruments, as
\begin{align}
\boldsymbol{Z}_{IVi}=\left(\boldsymbol{M}_{F^{0}_{X}}\boldsymbol{x}_{i}, \boldsymbol{M}_{F^{0}_{x,-1}}\boldsymbol{X}_{i,-1}  \right)
\end{align}

The first step IV estimator can be expressed as
\begin{align}
\begin{split}
\boldsymbol{\hat{\varphi}}_{IVi}&=\left(\left(\frac{\boldsymbol{Z}^{'}_{i}\boldsymbol{M}_{F^{0}_{X}}\boldsymbol{W}_{i}}{T } \right)^{'} \left(\frac{\boldsymbol{Z}^{'}_{i}\boldsymbol{M}_{F^{0}_{X}}\boldsymbol{Z}_{i}}{T}  \right)^{-1}  \left(\frac{\boldsymbol{Z}^{'}_{i}\boldsymbol{M}_{F^{0}_{X}}\boldsymbol{W}_{i}}{T}  \right)\right)^{-1} \\
&\left(\left( \frac{\boldsymbol{Z}^{'}_{i}\boldsymbol{M}_{F^{0}_{X}}\boldsymbol{W}_{i}}{T}  \right)^{'} \left(\frac{\boldsymbol{Z}^{'}_{i}\boldsymbol{M}_{F^{0}_{X}}\boldsymbol{Z}_{i}}{T}  \right)^{-1}  \left(\frac{\boldsymbol{Z}^{'}_{i}\boldsymbol{M}_{F^{0}_{X}}\boldsymbol{y}_{i}}{T}  \right)\right).
\end{split}
\end{align}

\section{Monte Carlo simulation design}
\subsection{dynamic heterogeneous panels data model without error factor structure}
The data generating process:
\begin{align}
\begin{split}
y_{i,t}&= \phi_{i} y_{i,t-1}+ \sum^{k}_{\ell=1}\beta_{\ell i}x_{\ell i,t}+u_{i,t}, \,\, for\,\,i=1,\ldots N;\,t=-49,\ldots,T\, , \\ \label{M1}
x_{\ell i,t}&=\sum^{k}_{\ell=1}\phi_{\ell i}x_{\ell i,t-1}+v_{\ell i,t},
\end{split}
\end{align}
where $u_{i,t}\overset{i.i.d.}{\sim} \mathcal{N}(0,\,1)\,,$ and $v_{\ell i,t}=\rho_{v, \ell}v_{\ell i,t-1}+\left( 1-\rho^{2}_{v, \ell} \right)^{\frac{1}{2}}\varpi_{\ell i,t}, \varpi_{\ell i,t} \overset{i.i.d.}{\sim} U(0.5, 1.5) \, ,\rho_{v, \ell}=0.5.$

The slope coefficients are generated as
\begin{align}
\phi_{i}=\phi+\eta_{\phi i},\,\, \beta_{1,i}=\beta_{1}+\eta_{\beta_{1} i}\, and\, \beta_{2,i}=\beta_{2}+\eta_{\beta_{2}i}.
\end{align}
Here we consider $\phi \in \left\{0.5 \right\}$, $\beta_{1}=3$ and $\beta_{2}=1$. For the design of heterogenous slopes, $\eta_{\phi i} \overset{i.i.d.}{\sim} U\left( -c, c\right)$, and
\begin{align}
\eta_{\beta_{\ell}i}= \left(1-\rho^{2}_{\beta}  \right)^{1/2}\eta_{\phi i}.
\end{align}
Here, we set $c=0.2,\, \rho_{\beta}=0.4$ for $\ell=1,2.$

\subsection{Dynamic heterogeneous panels data model with multi-factor error structure}
This Monte Carlo simulation design same as \citet{Norkute:2019}.
For convenience, we rewrite the data generating process as bellow
\begin{align}
y_{i,t}=\alpha_{i}+\phi_{i} y_{i,t-1}+ \sum^{k}_{\ell=1}\beta_{\ell i}x_{\ell i,t}+u_{i,t}, \,\, for\,\,i=1,\ldots N;\,t=-49,\ldots,T\, . \\ \label{M1}
\end{align}
We allow error factor structure in the model as
\begin{align}
u_{i,t}=    \sum^{m_{y}}_{s=1}\gamma^{0}_{si}f^{0}_{s,t} +\varepsilon_{i,t},
\end{align}
where
\begin{align}
f^{0}_{s,t}=\rho^{0}_{s,t} f^{0}_{s,t-1}+\left( 1-\rho^{2}_{fs} \right)^{1/2}\zeta_{s,t},
\end{align}
with $\zeta_{s,t} \overset{i.i.d.}{\sim} N(0,1)$ for $s=1,\ldots m_{y}$. We assume $k=2$ and $m_{y}=1+k=3$ and set $\rho^{0}_{s,t}=0.5$ for all $s$. The error term, $\varepsilon_{i,t}$, setting as
\begin{align}
\varepsilon_{i,t}=\varsigma_{\varepsilon}\sigma_{it}\left(\epsilon_{it}-1 \right)/\sqrt{2},
\end{align}
where $ \epsilon_{it}\overset{i.i.d.}{\sim} \chi^{2}_{1},\, \sigma^{2}_{it}=\eta_{i}\varphi_{t}, \,\eta_{i}\overset{i.i.d.}{\sim}\chi^{2}_{2}/2,$  and  $\varphi_{t}=t/T\, for \,\, t=0,\ldots, T.$ And we set
\begin{align}
\varsigma_{\varepsilon}=\frac{\pi_{\mu}}{1-\pi_{\mu}}m_{y}.
\end{align}
we set $\pi_{\mu} \in \left\{ 3/4\right\}.$

The process of regressors is
\begin{align}
x_{\ell it}=\mu_{\ell i}++\sum^{k}_{\ell=1}\phi_{\ell i}x_{\ell i,t-1}+\sum^{m_{x}}_{s=1}\gamma^{0}_{\ell si}f^{0}_{s,t}+v_{\ell it}, \,\, for\,\,i=1,\ldots N;\,t=-49,\ldots,T\,; \, \ell=1,2.
\end{align}
We set number of factor, $m_{x}$, is $2$. Therefore, $\boldsymbol{f}^{0}_{y,t}=\left(f^{0}_{1t}, f^{0}_{2t},f^{0}_{3t}  \right)^{'}$ and $\boldsymbol{f}^{0}_{x,t}=\left(f^{0}_{1t}, f^{0}_{2t}  \right)^{'}.$
We set
\begin{align}
v_{\ell i,t}=\rho_{v, \ell}v_{\ell i,t-1}+\left( 1-\rho^{2}_{v, \ell} \right)^{\frac{1}{2}}\varpi_{\ell i,t},  for \, \ell=1,2,
\end{align}
  where $\rho_{v, \ell}=0.5$ for all $\ell$.
The individual effect is
\begin{align}
 \alpha^{\ast}_{i}\overset{i.i.d.}{\sim} N\left(0, (1-\rho_{i})^{2} \right), \,\, \mu^{\ast}_{\ell i}=\rho_{\mu,\ell}\alpha^{\ast}_{i}+\left( 1-\rho^{2}_{\mu,\ell} \right)^{1/2}\omega_{\ell i},
\end{align}
where $\omega \overset{i.i.d.}{\sim}N\left(0, (1-\rho_{i})^{2} \right)$ and $\rho_{\mu,\ell}=0.5$.

Now, we define the factor loading in $u_{i,t}$ are generated as $\gamma^{0\ast}_{si}\overset{i.i.d.}{\sim}N\left(0, 1 \right)$, for $s=1,\ldots, m_{y}=3$, and the factor loading in $x_{1it}$ and $x_{2it}$ are drawn as
\begin{align}
\begin{split}
\gamma^{0\ast}_{1si}&=\rho_{\gamma,1s} \gamma^{0\ast}_{3i}+\left(1-\rho^{2}_{\gamma,1s}  \right)^{1/2}\xi_{1si};\,\xi_{1si}\overset{i.i.d.}{\sim}N\left(0, 1 \right); \\
\gamma^{0\ast}_{2si}&=\rho_{\gamma,2s} \gamma^{0\ast}_{si}+\left(1-\rho^{2}_{\gamma,2s}  \right)^{1/2}\xi_{2si};\,\xi_{2si}\overset{i.i.d.}{\sim}N\left(0, 1 \right);
\end{split}
\end{align}
for $s=1,\ldots, m_{x}=2$. We set $\rho_{\gamma, 11}=\rho_{\gamma, 12} \in \left\{ 0.5 \right\}$ and $\rho_{\gamma, 21}=\rho_{\gamma, 22}=0.5.$
The factor loading are generated as
\begin{align}
\boldsymbol{\Gamma}=\boldsymbol{\Gamma}^{0}+\boldsymbol{\Gamma}^{0\ast}_{i}
\end{align}
where
\begin{align}
\boldsymbol{\Gamma}^{0}_{i}=
\begin{bmatrix}
\gamma^{0}_{1i} & \gamma^{0}_{11i} & \gamma^{0}_{21i} \\
\gamma^{0}_{2i} & \gamma^{0}_{12i} & \gamma^{0}_{22i} \\
\gamma^{0}_{3i} &  0               &     0
\end{bmatrix}
\end{align}
and
\begin{align}
\boldsymbol{\Gamma}^{0\ast}_{i}=
\begin{bmatrix}
\gamma^{0\ast}_{1i} & \gamma^{0\ast}_{11i} & \gamma^{0\ast}_{21i} \\
\gamma^{0\ast}_{2i} & \gamma^{0\ast}_{12i} & \gamma^{0\ast}_{22i} \\
\gamma^{0\ast}_{3i} &  0               &     0
\end{bmatrix}.
\end{align}
We set
\begin{align}
\boldsymbol{\Gamma}^{0}=
\begin{bmatrix}
1/4 & 1/4 & -1 \\
1/2 & -1  & 1/4 \\
1/2 & 0   & 0
\end{bmatrix} .
\end{align}
And
\begin{align}
\alpha_{i}=\alpha+ \alpha^{\ast}_{i}, \, \mu_{\ell i}= \mu_{\ell}+\mu^{\ast}_{\ell i},
\end{align}
where $\alpha=1/2$, $\mu_{1}=1$, $\mu_{2}=-1/2$.

The slope coefficients are generated as
\begin{align}
\phi_{i}=\phi+\eta_{\phi i},\,\, \beta_{1,i}=\beta_{1}+\eta_{\beta_{1} i}\, and\, \beta_{2,i}=\beta_{2}+\eta_{\beta_{2}i}.
\end{align}
Here we consider $\phi \in \left\{0.5\right\}$, $\beta_{1}=3$ and $\beta_{2}=1$. For the design of heterogenous slopes, $\eta_{\phi i} \overset{i.i.d.}{\sim} U\left( -c, c\right)$, and
\begin{align}
\eta_{\beta_{\ell}i}=\left[(2c)^{2}/12 \right]\rho_{\beta}\xi_{\beta \ell i}+ \left(1-\rho^{2}_{\beta}  \right)^{1/2}\eta_{\phi i},
\end{align}
where
\begin{align}
\xi_{\beta \ell i}=\frac{\bar{v^{2}_{\ell i}}- \bar{v^{2}_{\ell}}}{\left[ N^{-1}\sum^{N}_{i=1} \left( \bar{v^{2}_{\ell i}}- \bar{v^{2}_{\ell }}\right)^{2} \right]^{1/2} },
\end{align}
with $\bar{v^{2}_{ell i}}=T^{-1}\sum^{T}_{t=1}v^{2}_{\ell i t}$, $\bar{v^{2}_{\ell}}=N^{-1} \sum^{N}_{i=1} \bar{v^{2}_{\ell i}}$, for $\ell=1,2.$
Here, we set $c=0.2,\, \rho_{\beta}=0.4$ for $\ell=1,2.$ And
\begin{align}
\varsigma^{2}_{v}=\varsigma^{2}_{\varepsilon}\left[SNR-\frac{\rho^{2}_{v}}{1-\rho^{2}_{v}}   \right]\left(\frac{\beta^{2}_{1}+\beta^{2}_{2}}{1-\rho^{2}_{v}}  \right)^{-1},
\end{align}

where $SNR=4$. For the $(T,N)$, we consider $T \in \left\{25, 50, 100, 200  \right\}$ and  $N \in \left\{25, 50, 100, 200  \right\}.$

\section{Monte Carlo simulation results}
\subsection{Dynamic Heterogeneous Panels without multifactor error structure }
We consider ARDL(1,0) model. \\
$\phi \in \left\{0.5 \right\}.$ \\
 $\beta_{1}=3.$ \\
  $\beta_{2}=1.$ \\
$u_{i,t}\overset{i.i.d.}{\sim} \mathcal{N}(0,\,1).$ \\
$\varpi_{\ell i,t} \overset{i.i.d.}{\sim} U(0.5, 1.5).$ \\
$\rho_{v, \ell}=0.5.$ \\
  $c=0.2.$ \\
  $ \rho_{\beta}=0.4.$ \\
 $T \in \left\{25, 50, 100, 200  \right\}.$ \\
 $N \in \left\{25, 50, 100, 200  \right\}.$ \\

LSMG estimator is provided in sheet 1 of MC.xlsx file. \\
IVMG estimator is provided in sheet 2 of MC.xlsx file. \\

\subsection{Dynamic Heterogeneous Panels with multifactor error structure }
We consider ARDL(1,0) model. \\
$\phi \in \left\{0.5 \right\}.$ \\
 $\beta_{1}=3.$ \\
  $\beta_{2}=1.$ \\
  $k=2.$ \\
  $m_{y}=1+k=3.$\\
  $m_{x}=k=2.$\\
$\zeta_{s,t} \overset{i.i.d.}{\sim} N(0,1)$ \\
$\pi_{\mu} \in \left\{ 3/4\right\}.$ \\
$\rho^{0}_{s,t}=0.5.$ \\
$\rho_{v, \ell}=0.5.$\\
 $\rho_{\mu,\ell}=0.5.$ \\
  $\gamma^{0\ast}_{si}\overset{i.i.d.}{\sim}N\left(0, 1 \right).$ \\
 $\xi_{1si}\overset{i.i.d.}{\sim}N\left(0, 1 \right).$ \\
$\xi_{2si}\overset{i.i.d.}{\sim}N\left(0, 1 \right).$ \\
$\rho_{\gamma, 11}=\rho_{\gamma, 12} \in \left\{ 0.5 \right\}.$ \\
$\rho_{\gamma, 21}=\rho_{\gamma, 22}=0.5.$ \\
$
\boldsymbol{\Gamma}^{0}=
\begin{bmatrix}
1/4 & 1/4 & -1 \\
1/2 & -1  & 1/4 \\
1/2 & 0   & 0
\end{bmatrix} .
$ \\
$\alpha=1/2.$ \\
 $\mu_{1}=1.$\\
  $\mu_{2}=-1/2.$ \\
$c=0.2.$ \\
$\rho_{\beta}=0.4.$ \\
$SNR=4.$ \\
$T \in \left\{25, 50, 100, 200  \right\}.$ \\
 $N \in \left\{25, 50, 100, 200  \right\}.$ \\

IVMG estimator is provided in sheet 3 of MC.xlsx file. \\

\section{Short summary}
\subsection{Dynamic Heterogeneous Panels without multifactor error structure}
1. The performance of IVMG estimator is better than LSMG estimator in bias and RMSE.

\subsection{Dynamic Heterogeneous Panels with multifactor error structure}
1. When $N$ and $T$ increase, the performance of IVMG estimator is good in bias and RMSE. \\
Related literature to dynamic Heterogeneous Panels with multifactor error structure: \citet{Chudik:2015} and \citet{Norkute:2019}.\\
Related literature to choosing number of instruments: \citet{Stephen:2001}, \citet{Swanson:2005}, \citet{Marine:2012}, \citet{Bai:2010} and \citet{Kang:2019}.

\newpage

\addcontentsline{toc}{section}{Reference}
\renewcommand\refname{References}
\bibliographystyle{chicago}
\bibliography{1}

\end{document}  \href{*}{*}